---
title: "WUP_04"
author: "Elizabeth Rivera"
date: "21 Sep 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
data = read.csv('GSScsv.csv')
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->


We try to predict WORDSUM based on degree. WORDSUM is numercial and DEGREE is categorical, therefore the method used will be used is numerical/ categorical.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
par(mar = c(5, 15, 5, 2))
boxplot(WORDSUM~DEGREE,data=data)
unique(data$DEGREE)
```


The x-axis represents the level of degree obtained and y-axis represents the average WORDSUM. Based on the graph the lowest average score on the WORDSUM test was obtained by LT HIGH SCHOOL scoring at 8. The highest average score on the WORDSUM test was obtained by the GRADUATES scoring at 11. Therefore, GRADUATES score higher than all other schooling DEGREES. 


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=data)
```

The numerical summary demonstrates the average score of GRADUATE performed the highest than all other schooling levels at 10.095238 and the lowest mean score is calculated at 7.750000 by the LT HIGH SCHOOL. 10.095238.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=data)
summary.aov(model1)
```


Based on the inferential results, the p-value is 0.000236 which is smaller than 0.05, therefore, we reject the null hypothesis. There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE.  



```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 17, 2, 2))
plot(TukeyHSD(model1),las=2)
```
Based on the above, it is noted that HIGH SCHOOL-GRADUATE and LT HIGH SCHOOL-HIGH SCHOOL's demonstrates there is a significant difference between their average WORDSUM score compared to the different level of schooling. Resulting, in lower family-wise confidence level for HIGH SCHOOL-GRADUATE and LT HIGH SCHOOL-HIGH SCHOOL level of schooling.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The study shows there is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE. The p-value is 0.000236 which is less than 0.05, resulting, we reject the null hypothesis. The data above demonstrates WORDSUM average score results are high in those in higher levels of education. 


